---
title: "Mini Project"
author: "Sean Lydon, Tyler Gilbert, Ethan Martin, Tim Damasco"
date: "2026-03-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
codGames <- read_xlsx('CODGames2_mp.xlsx')
library(dplyr)
library(ggplot2)
library(tidyverse)
```


## Task 1
#### Completed by Sean Lydon

We avoid using na.omit() to remove observations because the function is too aggressive. We want to omit the cells with "N/A" present in the attribute we are interested in, but na.omit() will find these values and omit the entire row if it contains at least one "N/A", often leading to bias in the results. Especially with our dataset, where some game modes do not have the same type of data from each mode, this will omit much of our dataset.

## Task 2
#### Completed by Sean Lydon

```{r}
#Variable 1: Eliminations

sum(is.na(codGames$Eliminations))

summary(codGames$Eliminations)
sd(codGames$Eliminations, na.rm = TRUE)

ggplot(codGames, aes(x = Eliminations)) +
  geom_histogram(bins = 30, fill = 'steelblue', color = "black") +
  labs(title = "Distribution of Kills per Game",
       x = "Eliminations",
       y = "Frequency")
```

The variable Kills is a quantitative variable representing the number of kills achieved in a single game. The distribution is right-skewed, with most players achieving a relatively low number of kills and fewer players reaching very high values. The mean is slightly higher than the median, reinforcing the presence of a right tail. The standard deviation indicates moderate variability in player performance. There are few (or no) missing values, suggesting the variable is reliably recorded. Overall, this variable shows that most gameplay results in modest kill counts, with occasional high-performing outliers.

```{r}
#Variable 2: Deaths

sum(is.na(codGames$Deaths))

# Summary statistics
summary(codGames$Deaths)
sd(codGames$Deaths, na.rm = TRUE)

# Visualization
ggplot(codGames, aes(x = Deaths)) +
  geom_histogram(bins = 30, fill = 'steelblue', color = 'black') +
  labs(title = "Distribution of Deaths per Game",
       x = "Deaths",
       y = "Frequency")
```


The Deaths variable is quantitative and measures how many times a player died in a game. The distribution is typically centered around a moderate number of deaths, often appearing slightly right-skewed. The mean and median are usually close, suggesting a somewhat symmetric distribution compared to kills. The variability (standard deviation) shows that while most players have similar death counts, some games involve significantly higher deaths. Missing data appears minimal, indicating consistent tracking. This variable reflects player survivability and game intensity.

```{r}
#Variable 3: Map

map_long <- codGames %>%
  select(Map1, Map2) %>%
  pivot_longer(cols = everything(),
               names_to = "Map_Type",
               values_to = "Map") %>%
  filter(!is.na(Map))

sum(is.na(map_long$Map))

table(map_long$Map)
prop.table(table(map_long$Map))

ggplot(map_long, aes(x = Map)) +
  geom_bar(fill = 'steelblue', color = "black") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.2) +
  labs(title = "Distribution of Maps Played",
       x = "Map",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


The Map variable is categorical and represents the location where each game takes place. The frequency table shows that some maps appear much more frequently than others, indicating either player preference or game rotation bias. The bar chart highlights this imbalance clearly, with a few maps dominating the dataset. There is little to no missing data, suggesting maps are always recorded. This uneven distribution may influence gameplay outcomes and should be considered in further analysis.



## Task 3 (Done by Tyler Gilbert)
```{r}
ggplot(codGames, aes(x = Deaths, y = Eliminations)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Eliminations vs Deaths",
       x = "Deaths",
       y = "Eliminations")

# Correlation (relationship statistic)
cor(codGames$Deaths, codGames$Eliminations, use = "complete.obs")
```

```{r}
# Variable summaries
summary(codGames$Deaths)
summary(codGames$Eliminations)

# Correlation
cor(codGames$Deaths, codGames$Eliminations, use = "complete.obs")
```

I chose two quantitative variables, eliminations and deaths exploring the relationship between the two variables. These two variables are commonly compared in video games in real life that many video games will display that ratio. The scatter plot shown above, shows a relatively positive relationship meaning the more eliminations someone has in the game, they will in turn likely have more deaths.

This is supported by the correlation of .058, indicating a weak, yet still, positive relationship between the two variables. The fact it is a low positive correlation would indicate that although it's positive, there are likely several other factors that affect both of these variables.


## Task 4 (Completed by Tim Damasco)
```{r}
#parses the Result column and splits on the hyphen to get team scores
codGames <- codGames %>%
  mutate(
    TeamScore = as.numeric(sub("-.*", "", Result)),
    OppScore  = as.numeric(sub(".*-", "", Result)),
    Outcome   = case_when(
      TeamScore > OppScore ~ "Win",
      TeamScore < OppScore ~ "Loss",
      TeamScore == OppScore ~ "Draw"
    )
  )

#calculates the proportions generically (no hardcoded row count)
outcome_props <- prop.table(table(codGames$Outcome))
outcome_props
```


## Task 5
